EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-08-17 draft.tex
Update draft.tex

% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
% This is done on the parent section, I renamed the sections to clarify this - M
To measure the speedup obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work, because we are using a stochastic search. 
%% JJ, can you please check if this is clear or if it needs more explanation?
We are going to measure the total increase of work done in a second by adding more workers.
The baseline for the comparison will be the algorithm with a single worker, 
and the minimum amount of work available will be five populations. 
This number is selected because the controller needs at least
three populations to migrate individuals between them, and we add two more to
keep even a single worker busy all the time. Figure \ref{fig:spworker:evalspersecond} 
shows the number of function evaluations per second, achieved by a certain number of
workers, for ten and twenty dimensions of $f_4$.

@@

%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{secs_per_instance}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes. The y-axis is shown in log scale.  }
  \label{fig:spworker:time}
\end{figure}
%
\begin{table}[tbp]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations} % time or rate? - JJ
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrrr}
    \hline
              &         & 5 Populations    & 10 Populations & 20 Populations \\
     Dimension& Workers & Median of rate & Median of rate & Median of rate  \\
    \hline
          10  & 1w      & 22.24  & 20.48           & 24.55  \\
              & 2w      & 10.31  & 10.00           & 12.79 \\
              & 4w      &  6.23  & 4.75            &  5.70 \\
              & 8w      &  5.86  &  2.33           & 2.74 \\
              & 16w     &  5.96  &  2.33           & \textbf{1.59}\\

@@

Figure \ref{fig:spworker:time} shows the execution time achieved by a certain number of
workers, for every problem dimension, which, besides, changes the
population size.

Each box-plot indicates the time required for the execution of 30 function
instances, in seconds. Each row has a different scale on the y-axis, to
compensate for the dimension increments. It is important to notice that the
execution time can also be affected by the algorithm finding a solution with an
error smaller than the threshold, before completing all the FEs.

In Table~\ref{tab:time} we can see that as expected, and in most cases, 
increasing the number of workers reduces the time
required for the execution of an instance, 
mainly when using the same parameters
(number of populations and number of workers), that is, performance
scales with the number of workers for a fixed problem and number of
sub-populations used.

>>>>

To measure the speedup obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work, because we are using a stochastic search. 
We are going to measure the total increase of work done in a second by adding more workers.
The baseline for the comparison will be the algorithm with a single worker, 
and the minimum amount of work available will be five populations. 
This number is selected because the controller needs at least
three populations to migrate individuals between them, and we add two more to
keep even a single worker busy all the time. Figure \ref{fig:spworker:evalspersecond} 
shows the number of function evaluations per second, achieved by a certain number of
workers, for ten and twenty dimensions of $f_4$.

@@

%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{secs_per_instance}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes. The y-axis is shown in log scale.  }
  \label{fig:spworker:time}
\end{figure}
%
\begin{table}[tbp]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations}
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrrr}
    \hline
              &         & 5 Populations    & 10 Populations & 20 Populations \\
     Dimension& Workers & Median of rate & Median of rate & Median of rate  \\
    \hline
          10  & 1w      & 22.24  & 20.48           & 24.55  \\
              & 2w      & 10.31  & 10.00           & 12.79 \\
              & 4w      &  6.23  & 4.75            &  5.70 \\
              & 8w      &  5.86  &  2.33           & 2.74 \\
              & 16w     &  5.96  &  2.33           & \textbf{1.59}\\

@@

Figure \ref{fig:spworker:time} shows the median (30 runs) of the execution time achieved by 2, 4 , 8, and 16  
workers, to finish an instance of the function $f_4$, for dimensions 10 and 20, with 5, 10, and 20
population sizes.

Each box-plot indicates the time required for the execution of 30 function
instances, in seconds. Each row has a different log scale on the y-axis, to
compensate for the dimension increments. It is important to notice that the
execution time can also be affected by the algorithm finding a solution with an
error smaller than the threshold, before completing all the FEs.

In Table~\ref{tab:time} we can see that as expected, and in all cases, 
increasing the number of workers reduces the time required for the execution of an instance, 
mainly when using the same parameters (number of populations and number of workers), that is, performance
scales with the number of workers for a fixed problem and number of
sub-populations used.
